---
title: "airtabler"
output: 
  html_document: 
    keep_md: yes
---
Provides access to the [Airtable API](http://airtable.com/api)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install

```{r install, eval=FALSE}
devtools::install_github("ecohealthalliance/airtabler")
```

## Setup
> After you've created and configured the schema of an Airtable base from the
graphical interface, your Airtable base will provide its own API to create,
read, update, and destroy records. -  [airtable.com/api](http://airtable.com/api) 

## Get and store the API key
Generate the airtable API key from your [Airtable account](http://airtable.com/account) page.

__airtabler__ functions will read the API key from
  environment variable `AIRTABLE_API_KEY`. To start R session with the
  initialized environvent variable create an `.Renviron` file in your home directory
  with a line like this:
  
`AIRTABLE_API_KEY=your_api_key_here`

To check where your home is, type `path.expand("~")` in your R console.


## Usage

Create airtable base object:

```{r}
library(airtabler)

TravelBucketList <- 
  airtable(
    base = "appIS8u9n73hzwE7R", 
    tables = c("Destinations", "Hotels", "Travel Partners")
  )
```

_Note that you should replace the Airtable base identifiers and `record_id`s when running the examples._

### Get records
Use select function to get all records:
```{r get_records}
hotels <- 
  TravelBucketList$Hotels$select()

knitr::kable(hotels[, c("id","Name", "Stars", "Price/night")], format = "markdown")
```

Filter records with formula (see [formula field reference ](https://support.airtable.com/hc/en-us/articles/203255215-Formula-Field-Reference)).

```{r get_records_formula}
hotels <- 
  TravelBucketList$Hotels$select(filterByFormula = " ({Avg Review} > 8.5)" )

knitr::kable(hotels[, c("id","Name", "Stars", "Avg Review", "Price/night")], format = "markdown")

```

Sort data with sort parameter:
```{r get_records_sort}
hotels <- 
  TravelBucketList$Hotels$select(sort = list(
    list(field="Avg Review", direction = "desc"),
    list(field="Price/night", direction = "asc")
  ))


knitr::kable(hotels[, c("id","Name", "Stars", "Avg Review", "Price/night")], format = "markdown")

```

### Using page size and offset

Define page size with `pageSize`:
```{r offset}

hotels <- TravelBucketList$Hotels$select(pageSize = 3)
nrow(hotels)


```

Continue at offset, returned by previous select:
```{r}
hotels <- TravelBucketList$Hotels$select(offset = get_offset(hotels))
nrow(hotels)
```


To fetch all rows (even > 100 records) use `select_all`. The `select_all` 
function will handle the offset and return the result as a single object.

```{r select_all}
hotels <- TravelBucketList$Hotels$select_all()
nrow(hotels)

```


Other optional arguments:

* __fields__ A list of fields to be returned (instead of all fields).
* __view__ The name or ID of the view, defined on the table.
* __maxRecord__ The maximum total number of records that will be returned.

### Retrieve a record
Add the `record_id` argument to get the details of a record:

```{r get_single}
radisson <- 
  
  TravelBucketList$Hotels$select(record_id = "recgKO7K15YyWEsdb")

str(radisson$fields, max.level = 1)
```

### Insert a record
Insert a new record with `insert` function (API returns all record data - including new record ID):
```{r insert}
record_data <- list(
  Name = "New hotel",
  `Price/night` = 200,
  Stars = "****",
  Amenities = c("Hiking", "Gym"),
  Notes = "Just a sample record.\nWith extra line in notes."
)

new_hotel <- 
  TravelBucketList$Hotels$insert(record_data)

cat("Inserted a record with ID=", new_hotel$id, sep = "")
```


### Update a record
Update the price of the new hotel (API returns all record data):
```{r update}
new_hotel <- 
  TravelBucketList$Hotels$update(
    record_id = new_hotel$id, 
    record_data = list(
      `Price/night` = 120,
      Notes = "Check out the price!!!"
    )
  )

cat("Updated a record with ID=", new_hotel$id, ". ", 
    "New price: ", new_hotel$fields$`Price/night`, sep = "")
```

### Delete a record
```{r delete}
TravelBucketList$Hotels$delete(new_hotel$id)
```


## Working with data frames

Standard Airtable API does not accept a table of records. 
Functions `insert` and `update` accept a data.frame and
execute transactions (call Airtable API) row by row.

Insert records with a data frame:
```{r data.frames.insert}
two_records <- 
  data.frame(
    Name = c("Sample1", "Sample2"),
    `Price/night` = c(150, 180),
    Stars = c("***", "****"),
    Amenities = I(list(c("Wifi", "Pool"), c("Spa", "Laundry"))),
    Notes = c("Foo", "Bar"),
    
    check.names = FALSE,
    stringsAsFactors = FALSE
  )

new_records <-
  TravelBucketList$Hotels$insert(two_records)

```

Update records with a data frame:
```{r data.frames.update}
# change records
record_ids <- sapply(new_records, function(x) x$id)
two_records$`Price/night` <- two_records$`Price/night` + 5
two_records$Stars <- "*****"


updated <- 
  TravelBucketList$Hotels$update(
    record_id = record_ids, 
    record_data = two_records)

```

Delete multiple records:
```{r data.frames.delete}
# delete new records
record_ids <- sapply(new_records, function(x) x$id)
deleted <- 
  TravelBucketList$Hotels$delete(record_ids)
```


## Programming with airtabler

While having all airtable base tables and functions in one object 
is handy in interactive mode, it is recommended to use primitive
functions for adding, reading, updating and deleting when programming
R packages:

```{r primitives}
travel_base <- "appIS8u9n73hzwE7R"

# read data
hotels <- air_select(travel_base, "Hotels")

# get one record
radisson <- air_select(travel_base, "Hotels", record_id = "recgKO7K15YyWEsdb")

# create
inserted <- air_insert(travel_base, "Hotels", record_data)

# update
updated <- air_update(travel_base, "Hotels", record_id = inserted$id, record_data)

# delete
deleted <- air_delete(travel_base, "Hotels", record_id = inserted$id)

```

