---
title: "airtabler"
output: 
  html_document: 
    keep_md: yes
---
Provides access to the [Airtable API](http://airtable.com/api)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install

```{r install, eval=FALSE}
devtools::install_github("bergant/airtabler")
```

## Setup
> After you've created and configured the schema of an Airtable base from the
graphical interface, your Airtable base will provide its own API to create,
read, update, and destroy records. -  [airtable.com/api](http://airtable.com/api) 

## Get and store the API key
Generate the airtable API key from your [Airtable account](http://airtable.com/account) page.

__airtabler__ functions will read the API key from
  environment variable `AIRTABLE_API_KEY`. To start R session with the
  initialized environvent variable create an `.Renviron` file in your R home
  with a line like this:
  
`AIRTABLE_API_KEY=your_api_key_here`

To check where your R home is, type `normalizePath("~")` in your R console.

## Usage

Create airtable base object:

```{r}
library(airtabler)

TravelBucketList <- 
  air_base(
    base = "appIS8u9n73hzwE7R", 
    tables = c("Destinations", "Hotels", "Travel Partners")
  )
```

_Note that you should replace the Airtable base identifiers and `record_id`s when running the examples._

### Get all records
```{r get_records}
library(airtabler)

# get data
hotels <- 
  TravelBucketList$Hotels$get()

knitr::kable(hotels[, c("id","Name", "Stars", "Price/night")], format = "markdown")

```

Optional arguments to `get` function:

* __view__ The name or ID of the view, defined on the table.
* __limit__ A limit on the number of records to be returned.
  Limit can range between 1 and 100.
* __offset__ Page offset returned by the previous list-records
  call. Note that this is represented by a record ID, not a numerical offset.
* __sortField__ The field name to use for sorting
* __sortDirection__ "asc" or "desc". The sort order in which the
  records will be returned. Defaults to asc.

### Retrieve a record
Add the `record_id` argument to get the details of a record:

```{r get_single}
radisson <- 
  TravelBucketList$Hotels$get(record_id = "recgKO7K15YyWEsdb")

str(radisson$fields, max.level = 1)
```

### Insert a record
Insert a new record with `insert` function (API returns all record data - including new record ID):
```{r insert}
record_data <- list(
  Name = "New hotel",
  `Price/night` = 200,
  Stars = "****",
  Amenities = c("Hiking", "Gym"),
  Notes = "Just a sample record.\nWith extra line in notes."
)

new_hotel <- 
  TravelBucketList$Hotels$insert(record_data)

cat("Inserted a record with ID=", new_hotel$id, sep = "")
```


### Update a record
Update the price of the new hotel (API returns all record data):
```{r update}
new_hotel <- 
  TravelBucketList$Hotels$update(
    record_id = new_hotel$id, 
    record_data = list(
      `Price/night` = 120,
      Notes = "Check out the price!!!"
    )
  )

cat("Updated a record with ID=", new_hotel$id, ". ", 
    "New price: ", new_hotel$fields$`Price/night`, sep = "")
```

### Delete a record
```{r delete}
TravelBucketList$Hotels$delete(new_hotel$id)
```


## Working with data frames

Standard Airtable API does not accept a table of records. 
Functions `insert` and `update` accept a data.frame and
execute transactions (call Airtable API) row by row.

Insert records with a data frame:
```{r data.frames.insert}
two_records <- 
  data.frame(
    Name = c("Sample1", "Sample2"),
    `Price/night` = c(150, 180),
    Stars = c("***", "****"),
    Amenities = I(list(c("Wifi", "Pool"), c("Spa", "Laundry"))),
    Notes = c("Foo", "Bar"),
    
    check.names = FALSE,
    stringsAsFactors = FALSE
  )

knitr::kable(two_records,format = "markdown")

new_records <-
  TravelBucketList$Hotels$insert(two_records)

```

Update records with a data frame:
```{r data.frames.update}
# change records
record_ids <- sapply(new_records, function(x) x$id)
two_records$`Price/night` <- two_records$`Price/night` + 5
two_records$Stars <- "*****"

knitr::kable(two_records,format = "markdown")


updated <- 
  TravelBucketList$Hotels$update(
    record_id = record_ids, 
    record_data = two_records)

```

Delete multiple records:
```{r data.frames.delete}
# delete new records
record_ids <- sapply(new_records, function(x) x$id)
deleted <- 
  TravelBucketList$Hotels$delete(record_ids)
```


