---
title: "Generate-Metadata-and-Backup"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Generate-Metadata-and-Backup}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(airtabler)
```

This vignette will demonstrate how to generate and update metadata for an Airtable
base using the [Airtable metadata api](https://airtable.com/developers/web/api/get-base-schema) then export the base to create an offline backup.  

*Note* Make sure that you have the [environment variable](https://github.com/gaborcsardi/dotenv) `AIRTABLE_API_KEY` set with your [Airtable API token](https://airtable.com/create/tokens). Make sure your token has schema read and write priveleges. 

## Creating metadata from schema 

`air_generate_metadata_from_api` creates a data frame that holds the structural metadata for your base. Structural metadata describes how entities relate to one another, in this case, how fields relate to tables within the base. 


```{r generate metadata}
# set your base id
base = "appVjIfAo8AJlfTkx"
## create a list
api_metadata <- air_generate_metadata_from_api(base = "appVjIfAo8AJlfTkx")
str(api_metadata)
```

## Adding a metadata table to the base

```{r add metadata table}
status <- air_create_metadata_table(base = base,meta_data = api_metadata)
```

## Create a descriptive metadata table for the base

The `description` table will describe who created the base, when, and why.
Descriptive metadata is important for long term storage and reuse. The default
terms in the `description` table match core DCMI terms. Additional terms can be
added.

```{r descriptive metadata}
## create a basic descriptive metadata table

description <- air_generate_base_description(title = "Base Title",
                              creator = "Collin Schwantes",
                              created = "2023-04-03",
                              description = "This base is an example for airtabler")


# We could have provided keywords as comma separated values but lets make things 
# more interesting by presenting them as a vector
description_with_keywords <- air_generate_base_description(title = "Base Title",
                              creator = "Collin Schwantes",
                              created = "2023-04-03",
                              description = "This base is an example for airtabler",
                              keywords = list(c("Example","R","Package","Airtable")))

```

## Add a description table to the base

```{r add descriptive metadata}

## add our vanilla description table - preferred way to work with description data
air_create_description_table(base = base,description = description)


## add our description with keywords and - we want keywords to be multiple select
## Note this is not the preferred method but it is possible

length(names(description_with_keywords))

create_choices <- function(x){
  
  choice_list <- list()
  
   choice_list$choices <- purrr::map(x,function(x_item){
             list(name = x_item)
    })

   return(list(choice_list)) # wrap in an extra list to 
}

keyword_options <- create_choices(description_with_keywords$keywords[[1]])

air_create_description_table(base = base,description = description_with_keywords,
                             type = c(rep("singleLineText",9),"multipleSelects"),
                             options = c(rep(NA,9),keyword_options))


```

## Updating metadata tables

Your data will change so your metadata should update as well.

### Structural Metadata
```{r update structural metadata}

# get your metadata from the api
 metadata <- air_generate_metadata_from_api(base = base)

# run the function
update_log  <- air_update_metadata_table(base,metadata)

# the update log provides an overview of records that were updated, inserted, or deleted
# and fields that were created in the event that the structure of your metadata table 
# changed.

```

### Descriptive Metadata



```{r update descriptive metadata}

base_description <- air_get_base_description_from_table(base = base,table_name = "Description",
                                                        field_names_to_snakecase = FALSE)

base_description$description <- "Keep on updating"

## since the field types are already established, its a little easier to add multipleSelect keywords
base_description$keywords[[1]] <- append("New Keyword",base_description$keywords[[1]])

base_description <- base_description |> 
  dplyr::select(-createdTime)

## if your base_description obj has a record id field, use that for the join. Default is title
air_update_description_table(base = base,description = base_description,table_name = "Description",join_field = "id")

```


## Using the metadata to create a backup

Now that you have your metadata tables setup, lets see how we can use them to
create backups. The workflow pulls data down from airtable into R where it can
be written to whatever format you like. There is a built in function for writing
to versioned to CSVs. This is the recommended workflow for creating backups.

### Pull all data into R 

```{r dump to R}

## pull down the metadata table from airtable 
metadata <- air_get_metadata_from_table(base = base,table_name = "Meta Data",add_id_field = FALSE)

airtable_base <- air_dump(base = base,metadata = metadata,description = base_description)

summary(airtable_base)

```

### Create a set of versioned CSVs

This function creates a folder with a unique ID based on a hash of the list
passed to table_list. If the data in your base do not change, then the hash
won't change and no new version of the data will be written. 

```{r dump to csv}

#  dump to csv
air_dump_to_csv(table_list = airtable_base)

# Make a change to the description table to show how hashing works
base_description <- air_get_base_description_from_table(base = base,table_name = "Description",
                                                        field_names_to_snakecase = FALSE)

## since the field types are already established, its a little easier to add multipleSelect keywords
base_description$keywords[[1]] <- append("How will the hash change?",base_description$keywords[[1]])

base_description <- base_description |> 
  dplyr::select(-createdTime)

## if your base_description obj has a record id field, use that for the join. Default is title
air_update_description_table(base = base,description = base_description,table_name = "Description",join_field = "id")

## pull the changed based down
airtable_base_updated <- air_dump(base = base,metadata = metadata,description = base_description)

## dump to csv
air_dump_to_csv(table_list = airtable_base_updated)
```

## Backup all bases in a workspace

*Note* This routine requires personal access tokens that can read and write 
data and schema to all bases in a workspace. 

Here we are going to take advantage of the [`list bases`](https://airtable.com/developers/web/api/list-bases) 
endpoint in the airtable API. It will list all bases that a token has access to.

```{r workspace backup}

# get all bases associated with token
bases <- air_list_bases()

# generate the description for our second
description_2 <- air_generate_base_description(title = bases$bases$name[[2]],creator = "Collin Schwantes",created = Sys.Date(),description = "A base to demo bulk backups")

description_log <- air_create_description_table(bases$bases$id[[2]],description = description_2)

metadata_2 <- air_generate_metadata_from_api(bases$bases$id[[2]])

metadata_log <- air_create_metadata_table(bases$bases$id[[2]],meta_data = metadata_2)

## add metadata and descriptive data to bases list

base_descriptions <- purrr::map(bases$bases$id,function(base){
  air_get_base_description_from_table(base,table_name = "Description")
})


base_metadata <- purrr::map(bases$bases$id,function(base){
  air_get_metadata_from_table(base,table_name = "Meta Data")
})


bases$bases$description <-  base_descriptions
bases$bases$metadata <-  base_metadata

base_df <- bases$bases

for(i in 1:nrow(base_df)){
  base_item <- base_df[i,]
  table_list <- air_dump(base = base_item$id ,metadata = base_item$metadata[[1]],description = base_item$description[[1]])
 
  output_dir <- sprintf("Airtabler Workspace/%s", base_item$name)
   air_dump_to_csv(table_list,output_dir = output_dir)
}


```

## CSVs to Frictionless

```{r csv dump to frictionless}

```

## Frictionless to deposits
